An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-21 Revision/response.tex

FilteredVamana achieves a good balance, staying under 30 ms while achieving Recall@10 > 0.97.
While StitchedVamana achieves extremely low latency (sub-5 ms) due to its stitched multi-label index structure, its fixed edge connectivity may limit effectiveness under highly uncorrelated label distributions. Nevertheless, it consistently supports 90\%+ Recall@10 across diverse datasets, as shown in our evaluations.

%%%%%%%%%%%%%%%%%%%%%% W4 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{weakness}
		\textit{The authors have reported the peak memory usage of index construction only.
			They are also recommended to report the peak memory usage of query processing.}
	\end{weakness}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion. We are deeply aware of this deficiency, so we added the search peak memory usage of each algorithm in \textcolor{blue}{Section 4.2.1} of attribute filtering and \textcolor{blue}{Section 4.3.1}  of range filtering, and carefully analyzed the reasons.

\textit{\textbf{ \textcolor{blue}{Section 4.2.1 Search peak memory.}}}

\textcolor{blue}{As shown in Figure, we can find that the peak memory usage of each algorithm during search does not differ significantly. Among them, CAPS and NHQ have relatively higher peak memory usage, while the database's peak memory usage is comparatively lower.}

\textit{\textbf{ \textcolor{blue}{Section 4.3.1 Search peak memory.}}}

	\textcolor{blue}{As for the Search peak memory usage, we can find from the Figure that it is basically consistent with the Build peak memory usage trend. The algorithm needs to load the complete index into memory during search. Therefore, algorithms with high building overhead, such as WinFilter, also have high search overhead; algorithms with low building overhead, such as Faiss, also have low search overhead. For Milvus, search does not require full index loading and there is additional metadata overhead when building, so the performance is different from the trend.}

@@

%%%%%%%%%%%%%%%%%%%%%% D1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{The authors have cited an experimental paper [49],
			but have not discussed about the main difference between their paper and [49].}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion! The paper [49] is a comprehensive survey and evaluation of graph structure approximate nearest neighbor search algorithms. Our paper extends this to a more complex "hybrid query" scenario and evaluates approximate search algorithms that combine vector similarity and attribute filtering/range filtering. Since hybrid queries are developed based on ANNs, we refer to the metrics and evaluation architectures of different methods proposed in this paper.

%%%%%%%%%%%%%%%%%%%%%% D2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{In Section 1.1, I could not find convincing references for the requirements in practical applications.
			[12] and [36] are research papers only, but not from real applications.}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion! We have carefully searched and found no papers describing the application of attribute distribution and attribute selectivity. However, the real-world application scenarios of attribute distribution and attribute selectivity are indeed very extensive. For attribute distribution, the label attributes of many recommended application scenarios in reality conform to the power-law distribution. For attribute selectivity: For example, if you need to query data in which an attribute accounts for a very small proportion of the total, choosing a more advantageous algorithm will greatly improve the search effect. There are also databases with actual applications that mention attribute selectivity, such as Milvus and AnalyticDB-V. For the two parts you suggested, we will add more comprehensive citations in the paper.

### 2025-07-21 VLDB2026/main.tex
1

	In recent years, hybrid query algorithms develop rapidly, giving rise to numerous attribute filtering ~\cite{NHQ,diskann} and range filtering algorithms \cite{serf,iRangeGraph}. In practical applications, the performance of hybrid query algorithms is influenced not only by unstructured data but also closely related to structured data. For attribute filtering algorithms, factors such as the number of attributes, attribute distribution \cite{UNG}, and attribute selectivity \cite{ACORN} significantly impact algorithm performance. As for range filtering algorithms, the size of the query range and the characteristics of different datasets  play an important role in determining algorithm performance.

@@

	\textit{\textbf{\textcolor{blue}{Large-scale Datasets}}}

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}

	\textcolor{blue}{
	We conducted experiments on the text2image-mix dataset. This dataset contains both image and text vectors. It is a more complex multimodal dataset. As shown in Figure~\ref{fig:TrinityExtension——Attribution}, although the queries are more complex, UNG and NHQ still perform the best. CAPS is the next best.
	Since the dataset mixes two modalities, the ground truth of the query set usually corresponds to the same modality. The nearest neighbors in the index are often from the same modality because they are closer in distance. Therefore, the search results on this dataset are similar to those in Figure~\ref{fig:exp_1_1_SingleLabel_1thread}. UNG and NHQ perform the best. CAPS clusters based on vector similarity. It actually clusters the vectors into the same modality. Then it uses attribute filtering to divide the data further. This allows CAPS to quickly find nearest neighbors that meet the filter conditions.
	In contrast, the text2image dataset shown in Figure~\ref{fig:exp_6_1} is different. The queries and the dataset are from different modalities. The queries are far from the dataset. The ground truths are also scattered and not clustered together. To find the ground truths, the algorithm often needs to traverse more points. This is a greater challenge for the algorithms.
%	We conducted extensive scalability experiments on the Text2Image-Mix dataset. This dataset contains both image and text vectors. It represents a more complex multimodal query scenario. As shown in Figure 12, the UNG and NHQ algorithms still achieve the best performance, followed closely by CAPS.
%	In the Text2Image-Mix dataset, the ground truth of the query vector is usually the corresponding vector in the same modality. Since the neighbor relationships in the graph are often based on vector distance, vectors of similar modalities naturally form closer connections in the graph. Therefore, during the search process, the algorithm tends to traverse efficiently within the same modality. This makes the search mechanism similar to our experiments on single-modality datasets, thus maintaining high performance.
%	The good performance of the CAPS algorithm is due to its unique hierarchical space partitioning strategy. It first clusters vectors based on similarity. In multimodal data, this clustering naturally groups vectors of similar modalities into the same cluster. Then, CAPS further partitions these clusters using attributes. This allows it to quickly find the nearest neighbors that meet the filtering conditions.
%	It is worth noting that, compared to the OOD dataset discussed in the "Dataset Impact" section, the Text2Image-Mix dataset also involves modality differences. However, its complexity lies in modality mixing. The OOD dataset, on the other hand, has a modality gap. This often requires traversing a broader and more sparse search space to find the ground truth. It poses a more severe challenge to the algorithm.
}

@@

	\textit{\textbf{\textcolor{blue}{Multi-modal Queries}}}
	\textcolor{blue}{
	Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure~\ref{fig:MultiModel range} is also consistent with the single-modality results in Figure~\ref{fig:exp_8_2}. As shown in Figure~\ref{fig:exp_8_3}, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets.}

>>>>

	In recent years, hybrid query algorithms develop rapidly, giving rise to numerous attribute filtering ~\cite{NHQ,diskann} and range filtering algorithms \cite{serf,iRangeGraph}. In practical applications, the performance of hybrid query algorithms is influenced not only by unstructured data but also closely related to structured data. For attribute filtering algorithms, factors such as the number of attributes, attribute distribution \cite{UNG}, and attribute  selectivity \textcolor{orange}{\cite{analyticdb,milvus}} significantly impact algorithm performance. As for range filtering algorithms, the size of the query range and the characteristics of different datasets  play an important role in determining algorithm performance.

@@

	\textit{\textbf{\textcolor{blue}{Complex Queries.}}}

@@

	\textit{\textbf{\textcolor{blue}{Complex Queries.}}}
